A Simplified Formulation for Molecular Dynamics with Rigid Bodies

### 2016-06-14 rigid_bodies.tex
Small changes in the introductory section

\title{A Simplified Approach for Molecular Dynamics with Rigid Bodies}

@@

\section{Introduction}
The advances in hardware along with the improvements in molecular dynamics algorithms and software has allowed to extend both the time and length scales attainable in atomistic simulations. This progress has resulted in an ambitious attempt to elucidate complex physical phenomena, such as the mechanism underlying molecular motors, the pathways for nanoparticle assembly, the interfacial phenomena, just to name a few. Because of its complexity and being computationally demanding, this challenging task requires strategies to further reduce the computational time.

One of the strategies commonly used to speed up the simulations is the suppression of the uninteresting degrees of freedom, as might be the case of the intramolecular vibrations. The rigid body molecular dynamics appears as the more suitable approach for rigid molecules. In contrast to the widely used iterative methods,\cite{Ryckaert1977,Andersen1983} one is able to make an implicit treatment of the holonomic constraints, which is computationally more efficient. In addition, and most important, it is possible to obtain a reversible and symplectic integrator in the microcanonical ensemble. The property of being symplectic implies the conservation of volume in phase space. This feature, together with time reversibility are mandatory for numerical integrators used in hybrid monte carlo (HMC) methods.\cite{Duane1987} In this sense, Miller \textit{et al}.\cite{Miller2002} have introduced a symplectic reversible integrator for the rotational motion in terms of the unit quaternion. This approach has been successfully used in pure molecular dynamics as well as in HMC methods to simulate a wide range of systems including the ubiquitous case of water \cite{Sakamaki2011,Reinhardt2012,Palmer2014,Gonzales2014}, ice water \cite{Geiger2014}, hydrates \cite{Tribello2009,Gorman2012}, diffusing gases in metal-organic frameworks \citep{Ghoufi2010}. It is worth noting that it also can be used for systems designed as a collection of interconnected rigid bodies, which is a coarse-graining strategy widely used for proteins \citep{Terada2003}, molecular machines \citep{Akimov2008,Konyukhov2010}, nanoparticles \cite{Knorowski2012,Patra2013}, etc.

>>>>

\title{A Simplified Formulation for Molecular Dynamics with Rigid Bodies}

\author{Ana J. Silveira}
\email{asilveira@plapiqui.edu.ar}
\affiliation{Chemical Engineering Department, Escola de Qu\'imica, Universidade Federal do Rio de Janeiro, Rio de Janeiro, RJ 21941-909, Brazil}
\affiliation{Planta Piloto de Ingenier\'ia Qu\'imica, PLAPIQUI, Universidad Nacional del Sur, Camino La Carrindanga Km 7-CC: 717, Bah\'ia Blanca, Argentina}

@@

\section{Introduction}

Improvements in molecular dynamics algorithms, along with advances in software and hardware, have allowed researchers to extend both the time and length scales attainable in their atomistic simulations. This progress has fomented the ambitious attempt to elucidate complex physical phenomena such as the mechanisms underlying molecular motors, pathways for nanoparticle assembly, and interfacial phenomena, to name just a few. Because of its complexity and large computational demand, such a challenging task requires strategies to enlarge the achievable scales even further.

One of the strategies commonly used to speed up MD simulations is the suppression of unimportant degrees of freedom, as might be the case of intramolecular vibrations. Widely used iterative methods such as SHAKE\cite{Ryckaert1977} and RATTLE\cite{Andersen1983} achieve

the dynamics of rigid bodies is able to keep this type of constraint implicitly, and thus emerges as a suitable approach in these cases.

one is able to make an implicit treatment of the holonomic constraints, which is computationally more efficient. In addition, and most important, it is possible to obtain a reversible and symplectic integrator in the microcanonical ensemble. The property of being symplectic implies the conservation of volume in phase space. This feature, together with time reversibility are mandatory for numerical integrators used in hybrid monte carlo (HMC) methods.\cite{Duane1987} In this sense, Miller \textit{et al}.\cite{Miller2002} have introduced a symplectic reversible integrator for the rotational motion in terms of the unit quaternion. This approach has been successfully used in pure molecular dynamics as well as in HMC methods to simulate a wide range of systems including the ubiquitous case of water \cite{Sakamaki2011,Reinhardt2012,Palmer2014,Gonzales2014}, ice water \cite{Geiger2014}, hydrates \cite{Tribello2009,Gorman2012}, diffusing gases in metal-organic frameworks \citep{Ghoufi2010}. It is worth noting that it also can be used for systems designed as a collection of interconnected rigid bodies, which is a coarse-graining strategy widely used for proteins \citep{Terada2003}, molecular machines \citep{Akimov2008,Konyukhov2010}, nanoparticles \cite{Knorowski2012,Patra2013}, etc.
